---
title: "ch10_testing_dummy_dat"
format: html
editor: visual
---

## CH 10 First Viagra statistics and how to *F*-ratio is calculated

Certain calculations about dummy.dat. See textbook page 404.

1 Load libraries

```{r}
library(tidyverse)
library(crayon)
```

2 Load dummy.dat file (libido stats with dummy groups dummy1 and dummy2.

```{r}
vigr <- read.delim(file <-  here::here("data/dummy.dat"), header=TRUE)
view(vigr)
```

3 Let's change dose to a factor

```{r}
vigr$dose<-factor(vigr$dose, levels = c(1:3), labels = c("Placebo", "Low
Dose", "High Dose"))
view(vigr)
```

3 We can run regression model for data with dummy variables in the same way as for any other kind of regression, by executing:

```{r}
vigrModel <- lm(libido~dummy1 + dummy2, data = vigr)
summary(vigrModel)
```

Notice that in summary that difference between the means of th high-dose group and the placebo group is: (5.0 - 2.2 = 2.8) and between low-dose group and the placebo group (3.2 - 2.2 = 1).

## Total sum of squares.

$$
SS_T = \sum_{i=1}^{N}(x_i-\bar{x})^2
$$

Let's calculate total variance and multiply it by n-1 (in this case it is 14)

```{r}
grand_variance <-   var(vigr$libido)
cat(yellow("Grand variance: "))
grand_variance
n <- nrow(vigr)
n <- n - 1
cat(yellow("n-1: "))
n
sst_2 <- grand_variance * n
cat(yellow("Total sum of squares, SST = Grand variance X (n-1) = "))
cat(green$bold(round(sst_2,2), '\n'))
```

## Model sum of squares

How much of this variation the regression model can explain? The mathematical expression:

$$
SS_M = \sum_{n=1}^{k}n_k(\bar{x}_k-\bar{x}_{grand})^2 
$$

Let's first calculate needed values:

```{r}
grand_mean = mean(vigr$libido)
group_means <- aggregate(vigr$libido, list(vigr$dose), FUN=mean)
#view(group_means)
placebo_mean = group_means[1,2]
lowdose_mean = group_means[2,2]
highdose_mean = group_means[3,2]
#all group sizes are 5
k1 <-  5
k2 <-  5
k3 <-  5
```

And now we are ready to calculate the Model sum of squares SSM:

```{r}
ssm  <- k1*(placebo_mean - grand_mean)^2 + k2*(lowdose_mean - grand_mean)^2 + k3*(highdose_mean - grand_mean)^2
cat(yellow("SSM = "))
cat(green$bold(round(ssm, 3)))
```

## Residual sum of squares

See page 410 of textbook:

SSR can be calculated by looking at the differences between the score obtained by a person and the mean of the group to which the person belongs.

$$
SS_R =\sum_{i=1}^{n}(x_{ik}-\bar{x_k})^2
$$

But we can use variances of each group and calculate simply:

$$
SS_R=\sum s_k^2(n_k-1)
$$

In order to use this equation, we have to first calculate group variances:

```{r}
group_variances <- aggregate(vigr$libido, list(vigr$dose), FUN=var)
#view(group_variances)
placebo_variance = group_variances[1,2]
lowdose_variance = group_variances[2,2]
highdose_variance = group_variances[3,2]
```

And now we can calculate Residual Sum of Squares (SSR) by (notice that degrees of freedom is n - 1:

```{r}
ssr = placebo_variance * (k1 - 1) + lowdose_variance * (k2 - 1) + highdose_variance * (k3 - 1)
cat(yellow("SSR = "))
cat(green$bold(round(ssr, 3)))
```

## Mean Squares

See textbook 411.

Mean squares for model can be calculated:

$$
MS_M=\frac{SS_M}{df_M}
$$

Here degrees of freedom is number of groups - 1 = 2

```{r}
msm <- ssm / 2
cat(yellow("MSM = "))
cat(green$bold(round(msm, 3)))
```

And mean squares of residuals can be calculated:

$$
MS_R=\frac{SS_R}{df_R}
$$

Here degrees of freedom is (total degrees of freedom) - (model degrees of freedom) = 14 - 2 = 12.

```{r}
msr <- ssr / 12
cat(yellow("MSR = "))
cat(green$bold(round(msr, 3)))
```

## The *F*-ratio

The *F*-ratio is a measure of the ratio of the variation explained by the model and variation explained by unsystematic factors. In other words it is the ratio of how **good** the model is against how **bad** it is. It can be calculated:

$$
\it{F}=\frac{MS_M}{MR_R}
$$

We have calculated these both values so we can simply get F-ratio:

```{r}
f_ratio <- msm / msr
cat(yellow$italic("F"),yellow("- ratio = "))
cat(green$bold(round(f_ratio, 2)))
```

We can also find out, what p-value of this F-ratio is. Or we can check out **critical values of F-statistics table** (in textbook)

```{r}
# 2 and 12 are the degrees of freedom from previous equation:
pf(f_ratio, 2, 12, lower.tail = FALSE)
```
