---
title: "CH7_dummy_contrasts"
author: "Petri Haavisto"
format: html
editor: visual
---

## How to create dummy variables for Categorical predictors

Check out textbook page 302

1 Load libraries:

```{r}
library(tidyverse)
```

2 Load data:

```{r}
gfr <- read.delim(file <-  here::here("data/GlastonburyFestivalRegression.dat"), header=TRUE)
```

3 View data

```{r}
view(gfr)
```

JUMP TO ALTERNATIVE, if you prefer manual method \*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*

4 Create dummy variables for music taste. But first make music a factor. Factors are used to categorize data.

```{r}
gfr$music <- factor(gfr$music)
contrasts(gfr$music) <- contr.treatment(4, base = 4)
```

5 Now music has a contrast attribute

```{r}
gfr$music
```

ALTERNATIVE STEPS 4 AND 5: \*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*

Set contrasts manually

```{r}
gfr$music <- factor(gfr$music)
crusty_v_NMA <- c(1,0,0,0)
indie_v_NMA <- c(0,1,0,0)
metal_v_NMA <- c(0,0,1,0)
contrasts(gfr$music) <- cbind(crusty_v_NMA, indie_v_NMA, metal_v_NMA)
```

And view:

```{r}
gfr$music
```

ALTERNATIVE STEPS END \*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*

6 Now we can check regression with dummy variables.

```{r}
glastonburyModel <- lm(change ~ music, data = gfr)
summary(glastonburyModel)
```

7 These means represent the average change in hygiene scores for the three groups.

```{r}
round(tapply(gfr$change, gfr$music, mean, na.rm=TRUE), 3)
```
