---
title: "Multinomial logistic regression Ch 8"
author: "Petri Haavisto"
format: html
editor: visual
---

## Multinomial logistic regression Chat-Up lines Chapter 8.9

Check out textbook 346 THIS HAS TO BE DONE AGAIN WITH UPDATED METHODS. DOES NOT PROVIDE RIGHT RESULTS AT THE MOMENT.

1 Load libraries

```{r}
library(conflicted) 
library(tidyverse)
library(mlogit)
conflict_prefer("filter", "dplyr")
conflict_prefer("lag", "dplyr")
```

2 Load data from dat-file

```{r}
chatData <- read.delim(file <- here::here("data/Chat-Up Lines.dat"), header = TRUE)
view(chatData)
```

3 You can check if categorical data is factor with is.factor function:

```{r}
is.factor(chatData$Success)
is.factor(chatData$Gender)
```

4 And if not, make them factors. And then run test above again.

```{r}
chatData$Success <- factor(chatData$Success)
chatData$Gender <- factor(chatData$Gender)
# Or you can use as.factor() function
#chatData$Gender <-  as.factor(chatData$Gender)
#chatData$Success <- as.factor(chatData$Success)
```

5 Let's make male the baseline gender

```{r}
chatData$Gender<-relevel(chatData$Gender, ref = 1)

```

5 We will restructure the current data so one person has 3 rows:

```{r}
#mlChat <- mlogit.data(chatData, choice = "Success", shape = "wide")
mlChat <- dfidx(chatData, shape = "wide", choice = "Success")
mlChat
```

6 Let's create a model

```{r}
ChatModel <- mlogit(Success ~ 1 | Good_Mate + Funny + Gender + Sex + Gender:Sex + Funny:Gender, data = mlChat, reference.level  = 3)
summary(ChatModel)
```

OBS! There is something wrong with model above! Needs further research!

But you can still check model's coefficients with:

```{r}
data.frame(exp(ChatModel$coefficients))
```

We can also check the base model:

```{r}
chatBase <- mlogit(Success ~ 1, data = mlChat, reference.level = 3)
summary(chatBase)
```

Confidence intervals:

```{r}
exp(confint(ChatModel))
```
