---
title: "Ch 8 Testing Assumptions with Penalty Kicks"
author: "Petri Haavisto"
format: html
editor: visual
---

## Penalty kicks and assumptions Chapter 8

Check textbook page 342

1 Load libraries

```{r}
library(tidyverse)
library(car)
```

2 Load dat-file

```{r}
penaltyData <- read.delim(file <- here::here("data/penalty.dat"), header = TRUE)
head(penaltyData)
```

3 Create a glm model (generalized linear model). Remember to factor categorical data.

```{r}
penaltyData$Scored <- factor(penaltyData$Scored)
penaltyModel.2 <- glm(Scored ~ Previous + PSWQ + Anxious, data = penaltyData, family = binomial())
summary(penaltyModel.2)
```

4 Now we can get VIF and tolerance (which is reciprocal of th VIF)

```{r}
vif(penaltyModel.2)
1/vif(penaltyModel.2)
```

5 Testing the linearity of the logit (textbook 344). OBS! here Log is natural log LN with Euler's number e.

```{r}
penaltyData$logPSWQInt <- log(penaltyData$PSWQ) * penaltyData$PSWQ
penaltyData$logAnxInt <- log(penaltyData$Anxious) * penaltyData$Anxious
penaltyData$logPrevInt <- log(penaltyData$Previous + 1) * penaltyData$Previous #missing data, add constant 1
view(penaltyData)
```

6 Let's add these three terms of each predictor and its log to the model

```{r}
penaltyTest.1 <- glm(Scored ~ PSWQ + Anxious + Previous + logPSWQInt + logAnxInt + logPrevInt, data = penaltyData, family = binomial())
summary(penaltyTest.1)
```
